Pre-/Post-Trikafta in CFRD

### 2022-04-22 Christine Chan/Pre-post Trikafta Prospective/pre_post_trikafta_prospective_analysis.Rmd
CGM pre/post trikafta data cleaning

```{r cgm data,include=FALSE}
# Merge
cgm = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/cgm_vars.csv")
cgm$Study.ID = str_extract(cgm$subject_id,"^[[:upper:]]{2}\\d{4}")
cgm$Event.Name = 
  str_extract(tolower(cgm$subject_id),"visit\\d{1}|visit \\d{1}|yr\\d{1}|year\\d{1}|year \\d{1}")
cgm$Event.Name = paste("Visit",gsub("[[:alpha:]]|\\s","",cgm$Event.Name))
df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
keep_a1c = df %>% group_by(Study.ID) %>% 
  summarise(n = sum(!is.na(A1C.result))) %>%
  filter(n == 2) %>% pull(Study.ID)
# Update DXA variables
dxa_vars = c("Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
            "BMD.Total.z.score","BMD.Subtotal.z.score",
            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score")
dxa = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/dxa_data.csv")
dxa = dxa %>% select(Study.ID,all_of(dxa_vars))
df = left_join(df,dxa)
keep_dxa = dxa$Study.ID
```

@@

```{r results='asis'}
t1 = tableby(~Age.at.visit.+Weight..kg.+weightz+trikafta_duration+
               Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
             data = df[df$Event.Name == "Visit 1" & df$Study.ID %in% keep_a1c,])
summary(t1,digits = 2,labelTranslations = newlabels)
```

@@

```{r results='asis'}
t2_vars = c("A1C.result","Weight..kg.","weightz","BMIz","FEV1..","FVC....",
            "trikafta_duration","Group","Glucose.0.minute",
            "Glucose.60.min","Glucose.120.min","c.peptide.0.minute",
            "`Glucose iAUC`","`C Peptide iAUC`","`Insulin iAUC`",
            "`C Peptide Index`","Matsuda",
            "`Adapted Matsuda`","`C Peptide ODI`","`ODI (Cpeptide index x Matsuda)`",
            "iAUC_ratio","homa2B","homa2S","homa2IR",
            "Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
            "BMD.Total.z.score","BMD.Subtotal.z.score",
            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score",
            "mean_sensor","max_sensor","min_sensor","percent_time_70_180",
            "percent_time_over_140","percent_time_over_180","percent_time_over_250",
            "percent_time_under_70","percent_time_under_54","standard_deviation","cv","r_mage")
f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

@@

```{r results='asis'}
f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
p <- paired(f,data = df[df$Study.ID %in% keep_a1c,], id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
            strata = `Age Group`)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

>>>>

```{r cgm data,include=FALSE}
# Merge
cgm = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/cgm_vars.csv")
cgm$Study.ID = str_extract(cgm$subject_id,"^[[:upper:]]{2}\\d{4}")
cgm$Event.Name = 
  str_extract(tolower(cgm$subject_id),"visit\\d{1}|visit \\d{1}|yr\\d{1}|year\\d{1}|year \\d{1}")
cgm$Event.Name = paste("Visit",gsub("[[:alpha:]]|\\s","",cgm$Event.Name))
df = left_join(df,cgm,by = c("Study.ID", "Event.Name"))
# Update DXA variables
dxa_vars = c("Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
            "BMD.Total.z.score","BMD.Subtotal.z.score",
            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score")
dxa = read.csv("./Christine Chan/Pre-post Trikafta Prospective/Data_Cleaned/dxa_data.csv")
dxa = dxa %>% select(Study.ID,all_of(dxa_vars))
df = left_join(df,dxa)
keep_dxa = dxa$Study.ID
```

@@

```{r results='asis'}
t1 = tableby(~Age.at.visit.+Weight..kg.+weightz+trikafta_duration+
               Sex..Male.1.+BMIz+FEV1..+FVC....+trikafta_duration+
               Genotype+Pancreatic.Status+Kalydeco+Orkambi+Symdeko,
             control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
             data = df[df$Event.Name == "Visit 1",])
summary(t1,digits = 2,labelTranslations = newlabels)
```

@@

```{r results='asis'}
t2_vars = c("A1C.result","Weight..kg.","weightz","BMIz","FEV1..","FVC....",
            "trikafta_duration","Group","Glucose.0.minute",
            "Glucose.60.min","Glucose.120.min","c.peptide.0.minute",
            "`Glucose iAUC`","`C Peptide iAUC`","`Insulin iAUC`",
            "`C Peptide Index`","Matsuda",
            "`Adapted Matsuda`","`C Peptide ODI`","`ODI (Cpeptide index x Matsuda)`",
            "iAUC_ratio","homa2B","homa2S","homa2IR",
            "Tissue..g.","Total.Fat.Mass....","Fat.Body.Mass..g.","Fat.Free.Mass..g.",
            "Lean.Body.Mass..g.","Total.Mass..kg.","Fat.Free.Mass..",
            "Fat.Mass.Index..FMI...","Lean.body.mass.fat.mass..LBM.FM.",
            "Fat.Free.Mass.Index...FFMI.","Lean.Body.Mass.Index..LBMI.",
            "BMD.Total.z.score","BMD.Subtotal.z.score",
            "DXA.FMI.calculated.z.score","DXA.LBMI.calculated.z.score",
            "mean_sensor","max_sensor","min_sensor","percent_time_70_180",
            "percent_time_over_140","percent_time_over_180","percent_time_over_250",
            "percent_time_under_70","percent_time_under_54","standard_deviation","cv","r_mage")
f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

@@

```{r results='asis'}
f = as.formula(paste0("Event.Name ~ ",paste0(t2_vars,collapse = "+")))
p <- paired(f,data = df, id = Study.ID,numeric.test = "signed.rank",signed.rank.exact = F,
            control = tableby.control(numeric.stats = c(c("Nmiss", "medianq1q3", "range"))),
            strata = `Age Group`)
summary(p,pfootnote = T,labelTranslations = newlabels)
```

### 2022-04-22 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_analysis.Rmd
CGM pre/post trikafta data cleaning

```{r include=FALSE}
# Import cleaned data from Casey
df = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
df$X = NULL
# Fill in MRN and CFF ID for everyone
df = df %>% 
  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
# Assume that blanks for pancreatic and G tube indicate "no"
df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
# Convert dates
date_vars = c("DOB","Start","Date")
df[,date_vars] = lapply(df[,date_vars], function(d){
  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
})
# Time from start and filtering
df = df %>% group_by(CFF.ID) %>% fill(Start,.direction = "downup")
df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
# Add UM (they did not provide dates, so need Time column)
um_a1cs = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
um_bmi = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
um_demo = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
um_ogtt = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
um_pfts = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
um = full_join(um_a1cs,um_bmi)
um = full_join(um,um_ogtt)
um = full_join(um,um_pfts)
um = left_join(um,um_demo)
um = um %>% filter(!is.na(CFRD.Status))
df = full_join(df,um)
# Filter and fill constants
df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
  ungroup() %>% filter(!is.na(Time),abs(Time) <= 365,!is.na(CFF.ID),
                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
df$Sex = as.factor(df$Sex)
levels(df$Sex) = c("F","M")
# BMI category
df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days")),df$Age,na.rm = T)/365.25
# Treat adults as age = 20
df$Age_for_BMI = ifelse(df$Age > 20,20,df$Age)
df$BMIz = y2z(df$BMI,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.bmi)
df$BMIz[is.infinite(df$BMIz)] = NA
df$BMI_group = apply(df,1,function(r){
  age = as.numeric(r["Age"])
  if(!is.na(age) & age > 19){
    bmi = as.numeric(r["BMI"])
    group = cut(bmi,c(-Inf,18.5,25,30,35,40,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)",
                   "Obese (Class II)","Obese (Class III+)"),right = F)
  } else if (!is.na(age) & age <= 19){
    z = as.numeric(r["BMIz"])
    perc = pnorm(z)
    group = cut(perc,c(-Inf,5,85,95,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)"),right = F)
  } else {group = NA}
  return(as.character(group))
})
df$BMI_group = as.factor(df$BMI_group)
df$BMI_group = ordered(df$BMI_group,c("Under Weight","Normal Weight","Overweight",
                                      "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
# PFTs
old <- c("Caucasian","Black","Other")
new <- c(1,2,5)
df$ppFEV = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FEV1)
df$ppFVC = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FVC)
# Before or after start
df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = F)
# Most recent before/after
most_recent = df %>% group_by(CFF.ID,Before_After) %>% slice_min(abs(Time))
```

>>>>

```{r include=FALSE}
# Import cleaned data from Casey
df = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data.csv",na.strings = "NA")
df$X = NULL
# Fill in MRN and CFF ID for everyone
df = df %>% 
  group_by(CFF.ID) %>% fill(MRN,.direction = "downup") %>% 
  group_by(MRN) %>% fill(CFF.ID,.direction = "downup")
# Assume that blanks for pancreatic and G tube indicate "no"
df$Hispanic.Latinx[is.na(df$Hispanic.Latinx)] = FALSE
df$Pancreatic.Status[is.na(df$Pancreatic.Status)] = "Sufficient"
df$G.tube.in.past.12.months[is.na(df$G.tube.in.past.12.months)] = "No"
# Convert dates
date_vars = c("DOB","Start","Date")
df[,date_vars] = lapply(df[,date_vars], function(d){
  parse_date_time(d,orders = c("mdy","ymd","dmy"),quiet = T)
})
# Time from start and filtering
df = df %>% group_by(CFF.ID) %>% fill(Start,.direction = "downup")
df = df %>% mutate(Time = as.numeric(difftime(Date,Start,units = "days"))) 
# Add UM (they did not provide dates, so need Time column)
um_a1cs = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_a1cs.csv",na.strings = "")
um_bmi = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_bmi.csv",na.strings = "")
um_demo = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_demo.csv",na.strings = "")
um_ogtt = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_ogtt.csv",na.strings = "")
um_pfts = read.csv("./Colleen Wood/Prepost Triakfta/Data_Cleaned/um_pfts.csv",na.strings = "")
um = full_join(um_a1cs,um_bmi)
um = full_join(um,um_ogtt)
um = full_join(um,um_pfts)
um = left_join(um,um_demo)
um = um %>% filter(!is.na(CFRD.Status))
df = full_join(df,um)
# Filter and fill constants
df = df %>% group_by(CFF.ID) %>% fill(DOB,Start:CFRD.Diagnosis.Date,.direction = "downup") %>%
  ungroup() %>% filter(!is.na(Time),abs(Time) <= 730,!is.na(CFF.ID),
                       (!is.na(CFRD.Diagnosis.Date) | !is.na(CFRD.Status) | site == "UW")) %>%
  select(-MRN) %>% select(CFF.ID,site,DOB,Start,everything()) %>% rename(Site = "site")
df$Sex = as.factor(df$Sex)
levels(df$Sex) = c("F","M")
# BMI category
df$Age = pmax(as.numeric(difftime(df$Date,df$DOB,units = "days")),df$Age,na.rm = T)/365.25
# Treat adults as age = 20
df$Age_for_BMI = ifelse(df$Age > 20,20,df$Age)
df$BMIz = y2z(df$BMI,x = df$Age_for_BMI,sex = df$Sex,ref=nl4.bmi)
df$BMIz[is.infinite(df$BMIz)] = NA
df$BMI_group = apply(df,1,function(r){
  age = as.numeric(r["Age"])
  if(!is.na(age) & age > 19){
    bmi = as.numeric(r["BMI"])
    group = cut(bmi,c(-Inf,18.5,25,30,35,40,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)",
                   "Obese (Class II)","Obese (Class III+)"),right = F)
  } else if (!is.na(age) & age <= 19){
    z = as.numeric(r["BMIz"])
    perc = pnorm(z)
    group = cut(perc,c(-Inf,5,85,95,Inf),
        labels = c("Under Weight","Normal Weight","Overweight","Obese (Class I)"),right = F)
  } else {group = NA}
  return(as.character(group))
})
df$BMI_group = as.factor(df$BMI_group)
df$BMI_group = ordered(df$BMI_group,c("Under Weight","Normal Weight","Overweight",
                                      "Obese (Class I)","Obese (Class II)","Obese (Class III+)"))
# PFTs
old <- c("Caucasian","Black","Other")
new <- c(1,2,5)
df$ppFEV = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FEV1)
df$ppFVC = pctpred_GLI(df$Age,df$Height/100,gender = as.numeric(df$Sex),
            ethnicity = new[match(df$Race, old)],
            FEV1 = df$FVC)
# Before or after start
df$Before_After = cut(df$Time,c(-Inf,0,Inf),labels = c("Before","After"),right = T)
# Sort
df = df %>% arrange(Site,CFF.ID,Date)
# Write sorted DF
write.csv(df, file = "./Colleen Wood/Prepost Triakfta/Data_Cleaned/analysis_data_with_UM.csv",
          row.names = F,na = "")
# Most recent before/after - limit to 2 years before start
# Want the value closest to start date before start, and most recent with respect to now from start
# (i.e. the longest time after start)
most_recent = df %>% group_by(CFF.ID,Site,Before_After) %>%
  mutate(time_new = max(Time) - Time) %>%
  arrange(time_new) %>% 
  summarise(BMIz = first(na.omit(BMIz)),ppFEV = first(na.omit(ppFEV)),
            ppFVC = first(na.omit(ppFVC))) %>%
  arrange(Site,CFF.ID,Before_After)
t = df %>% select(CFF.ID,Site,Date,Time,Before_After,BMIz,ppFEV,ppFVC) %>%
  group_by(CFF.ID,Site,Before_After) %>%
  arrange(Site,CFF.ID,Before_After)
keep = most_recent %>% group_by(CFF.ID) %>% count() %>% filter(n == 2)
```

# Participant Characteristics

### 2022-04-22 Colleen Wood/Pre-post Trikafta/pre_post_trikafta_merge.Rmd
CGM pre/post trikafta data cleaning

NOTE: I HAVE THE SHARED DRIVE AS S:, to change if mapped to another letter
```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Christine Chan"
}
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# bmi, pulm function 577 obs
# 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
chco_bmi_pft = read_csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")

# continuous glucose monitoring, 23 obs
# 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
chco_cgm = read_csv("./Prepost Triakfta/Data_Cleaned/chco_cgm.csv")

# chco glycemic, 261 obs
# 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
chco_glycemic = read_csv("./Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")

# chco registry 108 obs
# CFF ID and MRN, start date
chco_registry = read_csv("./Prepost Triakfta/Data_Cleaned/chco_registry.csv")
###################################

@@

# cgm 16 obs
montana_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_cgm.csv")

@@

# glycemic 279
montana_glycemic_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv") # looks identical
###################################

@@

Lets try and stitch these sets together by site (one site at a time as above)
note: by default, the dplyr join functions join by all variables in common. So we need to make sure the common variables are same format and name
```{r form stitching}
####### CHCO ############
# CGM Data Cleaning
# get a CFF ID and visit variable by splitting subjectid variable into CFF ID and Visit
chco_cgm$`CFF ID` = as.numeric(sub("_.*", "", chco_cgm$subject_id))
chco_cgm$Visit = sub(".*_", "", chco_cgm$subject_id)

# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
# weird because one obs isnt in a date time, just a date so ill just brute it
chco_cgm = chco_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
                               Date = ifelse(datewithtime,sub(" .*", "", chco_cgm$date_cgm_placement), date_cgm_placement))

@@

chco_long = full_join( chco_registry, chco_long) # joins by MRN, CFFID, Start
chco_long = full_join(chco_long, chco_cgm) # joins by cff id and date

@@

# keep original data, add in a new one to match CHCO
montana_demo = montana_demo %>% rename(race_original = Race)
montana_demo = montana_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# rename the bmi_pft to chco names
montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
                                             FEV1 = `Measure of FEV1`,
                                             Height = `Patients height`,
                                             Weight = `Weight of Patient`)

# make registry dataframe
montana_registry = full_join(montana_demo, montana_genotypes)
montana_registry = full_join(montana_registry, montana_pancreatic)
montana_registry = full_join(montana_registry, montana_gtubes)

@@

# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
montana_cgm = montana_cgm %>% mutate(datewithtime = str_detect(date_cgm_placement, " "),
                               Date = ifelse(datewithtime,sub(" .*", "", montana_cgm$date_cgm_placement), date_cgm_placement))

@@

# combine forms
uw_long = full_join(uw_glycemic, uw_bmi_pft) # will join by cffid, Date
uw_long = full_join(uw_registry, uw_long) # joins by MRN, CFFID, Start
uw_long = full_join(uw_long, uw_cgm) # joins by cff id and date

# add in a site id variable
uw_long$site = "UW"
########################
```
TO ASK TIM: 
-significance of dates? adding a ton of different rows, and wondering which dates are significant
-why do UW and Montana files (pancreatic) just have dates? are these different dates/ significant? do we need this file/ rename the date? gtubes similar

>>>>

NOTE: I HAVE THE SHARED DRIVE AS S:, to change if mapped to another letter
```{r setup, include=FALSE}
library(arsenal)
library(tidyverse)
library(performance)
library(knitr)
library(lubridate)
library(stringr)
knitr::opts_chunk$set(echo = FALSE)
if(Sys.info()["sysname"] == "Windows"){
  home_dir = "S:/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
} else if (Sys.info()["sysname"] == "Linux"){
  home_dir = "~/UCD/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
} else if (Sys.info()["sysname"] == "Darwin"){
  home_dir = "/Volumes/PEDS/RI Biostatistics Core/Shared/Shared Projects/Vigers/CF/Colleen Wood"
}
knitr::opts_knit$set(root.dir = home_dir)
```

@@

# bmi, pulm function 577 obs
# 8 vars MRN - ID, date for visits; start - ? start of trt?, fev, bmi, height, weight, fec 
chco_bmi_pft = read_csv("./Prepost Triakfta/Data_Cleaned/chco_bmi_pft.csv")
chco_bmi_pft$Date = mdy(chco_bmi_pft$Date) 
# continuous glucose monitoring, 23 obs
# 105 vars subjectid - number_pre/post for visit and a date; numbers dont seem to correspond to mrn in bmi pft
chco_cgm = read_csv("./Prepost Triakfta/Data_Cleaned/chco_cgm.csv")
chco_cgm$Date = date(parse_date_time(chco_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))
# chco glycemic, 261 obs
# 7 var MRN and CFF id (=subjectid?), date for visits; start... etc
chco_glycemic = read_csv("./Prepost Triakfta/Data_Cleaned/chco_glycemic.csv")
chco_glycemic$Date = mdy(chco_glycemic$Date)
# chco registry 108 obs
# CFF ID and MRN, start date
chco_registry = read_csv("./Prepost Triakfta/Data_Cleaned/chco_registry.csv")
###################################

@@

# cgm 16 obs
montana_cgm <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_cgm.csv")
montana_cgm$Date = date(parse_date_time(montana_cgm$date_cgm_placement,
                                                orders = c("mdyHM","mdy")))

@@

# glycemic 279
montana_glycemic_registry <- read_csv("./Prepost Triakfta/Data_Cleaned/montana_glycemic_registry.csv")
montana_glycemic_registry$Date = mdy(montana_glycemic_registry$Date)
# looks identical
###################################

@@

Lets try and stitch these sets together by site (one site at a time as above)
note: by default, the dplyr join functions join by all variables in common. So we need to make sure the common variables are same format and name

```{r form stitching}
####### CHCO ############
# CGM Data Cleaning
# get a CFF ID and visit variable by splitting subjectid variable into CFF ID and Visit
chco_cgm$`CFF ID` = as.numeric(sub("_.*", "", chco_cgm$subject_id))
chco_cgm$Visit = sub(".*_", "", chco_cgm$subject_id)

@@

chco_long = full_join(chco_registry, chco_long) # joins by MRN, CFFID, Start
chco_long = full_join(chco_long, chco_cgm) # joins by cff id and date

@@

# keep original data, add in a new one to match CHCO
montana_demo = montana_demo %>% rename(race_original = Race)
montana_demo = montana_demo %>% mutate(Race = case_when(race_original == "White" ~ "Caucasian",
                                                        race_original == "Black or African American" ~ "Black",
                                                        race_original == NA ~ "Unknown",
                                                        TRUE ~ "Other"))
# rename the bmi_pft to chco names
montana_bmi_pft = montana_bmi_pft %>% rename(FVC = `Measure of FVC`,
                                             FEV1 = `Measure of FEV1`,
                                             Height = `Patients height`,
                                             Weight = `Weight of Patient`)
montana_bmi_pft$Date = mdy(montana_bmi_pft$Date)

# make registry dataframe
montana_registry = full_join(montana_demo, montana_genotypes)
montana_registry = full_join(montana_registry, montana_pancreatic)
montana_registry = full_join(montana_registry, montana_gtubes)
montana_registry$Date = mdy(montana_registry$Date)

@@

# Get a Date variable by taking the date from "date time" char variable date_cgm_placement
montana_cgm$Date = mdy(montana_cgm$Date)

@@

# combine forms
uw_long = full_join(uw_glycemic, uw_bmi_pft) # will join by cffid, Date
uw_long = full_join(uw_registry, uw_long) # joins by MRN, CFFID, Start
uw_long = full_join(uw_long, uw_cgm) # joins by cff id and date
uw_long$Date = mdy(uw_long$Date)
# add in a site id variable
uw_long$site = "UW"
########################
```

TO ASK TIM: 
-significance of dates? adding a ton of different rows, and wondering which dates are significant
-why do UW and Montana files (pancreatic) just have dates? are these different dates/ significant? do we need this file/ rename the date? gtubes similar
